Proof nets for first-order additive linear logic

### 2019-04-22 7 - FSCD 2019 - Final/ALL1.tex
one exists missing
one all1 in the appendix

%We investigate the \emph{proof-net problem}, finding a canonical representation of proofs, for additive linear logic (\all) extended with first-order quantification (\all1). \all\ is the fragment of linear logic \cite{Girard-1987} that describes sums and products, and their canonical morphisms: projections, injections, and diagonals. It is rich, well-behaved, and compact: an ideal microcosm in which to explore semantic and proof-theoretic ideas. Here, we explore two techniques to account for the witness information that is essential to first-order proof: \emph{explicit substitution} and first-order \emph{unification}.

@@

Here, we extend proof nets to first-order additive linear logic (\all1). Beyond the solution to the proof-net problem, a main contribution is the (further) development of the two techniques we consider: \emph{explicit substitutions} for witness assignment, and reconstruction of witness information through \emph{unification}, as pioneered for MLL by the second author \cite{Hughes-2018}. We expect to apply these to first-order logics more generally.

@@

\defn{Witness nets} record witness assignment in substitution maps attached to each link. The example below left shows the proof net for the sequent proof earlier. (We will assume a different variable for each quantifier, and we attach links to predicate letters, as the root connective of an atomic proposition.) Witness nets are canonical for \all1 sequent calculus permutations. Composition is direct, where the witness assignments of links are composed through a simple process of \emph{interaction\,+\,hiding} similar to that of game semantics \cite{Mimram-2011}.

@@

@@

% .................................................
\setcounterref{retheorem}{thm:proof->net->proof}
\addtocounter{retheorem}{-1}
\begin{retheorem}
For any \all1 proof $\pi$, the witness net $[\pi]$ sequentializes to $\pi$. 
\end{retheorem}

>>>>

%We investigate the \emph{proof-net problem}, finding a canonical representation of proofs, for additive linear logic (\all) extended with first-order quantification (\all{\textnormal 1}). \all\ is the fragment of linear logic \cite{Girard-1987} that describes sums and products, and their canonical morphisms: projections, injections, and diagonals. It is rich, well-behaved, and compact: an ideal microcosm in which to explore semantic and proof-theoretic ideas. Here, we explore two techniques to account for the witness information that is essential to first-order proof: \emph{explicit substitution} and first-order \emph{unification}.

@@

Here, we extend proof nets to first-order additive linear logic (\all{\textnormal 1}). Beyond the solution to the proof-net problem, a main contribution is the (further) development of the two techniques we consider: \emph{explicit substitutions} for witness assignment, and reconstruction of witness information through \emph{unification}, as pioneered for MLL by the second author \cite{Hughes-2018}. We expect to apply these to first-order logics more generally.

@@

\defn{Witness nets} record witness assignment in substitution maps attached to each link. The example below left shows the proof net for the sequent proof earlier. (We will assume a different variable for each quantifier, and we attach links to predicate letters, as the root connective of an atomic proposition.) Witness nets are canonical for \all{\textnormal 1} sequent calculus permutations. Composition is direct, where the witness assignments of links are composed through a simple process of \emph{interaction\,+\,hiding} similar to that of game semantics \cite{Mimram-2011}.

@@

	\infer[\!\!\scriptstyle{\mathrm{cut}}]{\seq{A\rho'}{\exists x.C\rho'}}{

@@

% .................................................
\setcounterref{retheorem}{thm:proof->net->proof}
\addtocounter{retheorem}{-1}
\begin{retheorem}
For any \all{\textnormal 1} proof $\pi$, the witness net $[\pi]$ sequentializes to $\pi$. 
\end{retheorem}
